\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{float}
\usepackage[skip=10pt, indent=20pt]{parskip}
\graphicspath{{./figures/}{./lessons/03-lcd/assets/docs/figures/}}

\title{Lesson 3 - The LCD Display}
\author{Bryan A. \textit{CrazyUncleBurton} Thompson \\ \textit{bryan@batee.com}}
\date{March 17, 2026}

\begin{document}

\maketitle


\section*{Overview}
This lesson will introduce outputting to the LCD display and reading from 
the touchscreen.
\vspace{3mm}


\section*{Theory}
We can output text and graphics to the LCD many different ways and discusss 
the advantages and disadvantages of each way.
\vspace{3mm}


\section*{Hardware Required}

The following hardware is needed for this lesson:

1 - M5Stack Tab5 Microcontroller Development Board

1 - M5Stack Unit ENV III Environmental Sensor

1 - 4 wire M5Stack sensor cable (comes with the above module)

1 - USB cable to connect the Tab5 USB-C port to your PC
\newpage


\section*{Objectives}
\begin{itemize}
  \item Draw static Text to the LCD using M5Unified.
  \item Draw static graphics to the LCD using M5GFX.
  \item Draw changing data to the LCD using M5Canvas.
  \item Read the Touchscreen Data.
  \item Build our Second Circuit.
  \item Build, Compile and Test the Programs.
  \item Discuss differences and advantages of the various modes presented.
\end{itemize}
\vspace{15mm}

\section*{The Tab5 LCD Display}

This is a 5" 1280x720 (720P) IPS Touchscreen display with an ST7123 Display 
Driver Chipset.

There are many ways that we can write to this display:
\begin{itemize}
  \item M5Unified - This is a supplied text library from M5Stack.  It is quick
  to learn, and is best at displaying static data to the LCD.  If the data 
  changes over time, this will result in flickering screens.
  \item M5GFX - This is a supplied graphics library from M5Stack.  It is best 
  at displaying static Text and graphics to the LCD.  If the data changes 
  content or position over time, this will result in flickering screens.
  \item M5Canvas - This is an extension of M5GFX which treats 
  the screen as a framebuffer or canvas or sprite.  We draw the next screen to 
  a temporary "canvas" or sprite in the background.  When the next screen has 
  been updated, we push the canvas to the display.  The result is changes
  without the flicker-free nature of the previous libraries.
  \item LVGL / Squareline Studio.  LVGL is a non-factory library, and 
  Squareline Studio is an app that lets us create GUIs.  We will present this 
  in the later GUI lesson.

\end{itemize}
\vspace{3mm}

\section*{The LCD Text Program}
\vspace{3mm}

\section*{LCD Text Connections}

\vspace{3mm}

\section*{Compiling and Uploading the LCD Text Program}
\vspace{3mm}
\begin{enumerate}
    \item See the upload port icon at the bottom of the screen. Ensure the 
    upload port is set to Auto. 
    \vspace{3mm}

    \item Make sure the \textbf{Switch PlatformIO Project Environment} icon 
    at the bottom of the screen is set to \textbf{3a-lcd-text}.
    \vspace{3mm}

    \item To compile the program and upload it to the microcontroller, click 
    on the \textbf{PlatformIO: Build} icon at the bottom of the screen.
    \vspace{3mm}

    \item The results of the compile and upload process are presented in the 
    terminal window at the bottom of the screen.
    \vspace{3mm}

\end{enumerate}
\vspace{3mm}

\section*{Running the LCD Text Program}

When the program uploads and the microcontroller reboots, we should see text 
appear on the LCD!
\newpage

\section*{The LCD Data Program}
\vspace{3mm}

\section*{LCD Data Connections}
\vspace{3mm}

\section*{Compiling and Uploading the LCD Data Program}
\vspace{3mm}
\begin{enumerate}
    \item See the upload port icon at the bottom of the screen. Ensure the 
    upload port is set to Auto. 
    \vspace{3mm}

    \item Make sure the \textbf{Switch PlatformIO Project Environment} icon 
    at the bottom of the screen is set to \textbf{3a-lcd-data}.  
    \vspace{3mm}

    \item To compile the program and upload it to the microcontroller, click 
    on the \textbf{PlatformIO: Build} icon at the bottom of the screen.
    \vspace{3mm}

    \item The results of the compile and upload process are presented in the 
    terminal window at the bottom of the screen.
    \vspace{3mm}

\end{enumerate}
\vspace{3mm}

\section*{Running the LCD Data Program}

When the program uploads and the microcontroller reboots, we should see data 
appear on the LCD, with smoothly changing screen updates!
\newpage

\section*{The LCD Graphics Program}
\vspace{3mm}

\section*{LCD Graphics Connections}
\vspace{3mm}

\section*{Compiling and Uploading the LCD Graphics Program}
\vspace{3mm}
\begin{enumerate}
    \item See the upload port icon at the bottom of the screen. Ensure the 
    upload port is set to Auto. 
    \vspace{3mm}

    \item Make sure the \textbf{Switch PlatformIO Project Environment} icon 
    at the bottom of the screen is set to \textbf{3a-lcd-graphics}.
    \vspace{3mm}

    \item To compile the program and upload it to the microcontroller, click 
    on the \textbf{PlatformIO: Build} icon at the bottom of the screen.
    \vspace{3mm}

    \item The results of the compile and upload process are presented in the 
    terminal window at the bottom of the screen.
    \vspace{3mm}

\end{enumerate}
\vspace{3mm}

\section*{Running the LCD Graphics Program}

When the program uploads and the microcontroller reboots, we should see 
graphics appear on the LCD.  At the bottom of the screen, a Next button 
lets us change to the next graphic screen.
\newpage


\section*{Conclusion}
This completes the LCD Display lesson.  We learned the following concepts:

\begin{itemize}
    \item Displaying static text on the LCD
    \item Displaying dynamic data on the LCD
    \item Displaying primitive graphics on the LCD
    \item Creating touch-sensitive soft buttons
\end{itemize}
\vspace{3mm}

In the next lesson, we will cover Digital Inputs and Outputs in much more 
detail than we have so far!

\end{document}
